Characterization of gaps in directly bonded Si compound optics

### 2014-04-15 AO_bondedSi.tex
more tweaks from DTJ up to page 10

While it is possible that proper precautions to eliminate all micron-sized particles from both Si surfaces prior to bonding, the power-law nature of the particle size distribution in clean-room environments \cite{doi:10.1080/02786828608959094} means that it is significantly more difficult to assure that no particles of the 10$-$60 nm size scale are present.  Small particles could be a concern if they act as "tent poles", that create gaps where the bonding is incomplete (Figure \ref{figParticle}).  Authors have referred to these large ($\sim$ 1 $\mu$m size) gaps as \emph{IR bubbles}, or simply bubbles, owing to their appearance in infrared (IR) transmission images as contiguous regions of a deficit of IR transmission.  They have also been referred to as voids, or interfacial voids, or gaps.  In this paper we refer to them as gaps.

@@

The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.  We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.  The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.  There is also evidence that bond front propagation is much slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.

@@

\begin{table}[h!]
\caption{UTexas Si bonding experiments \label{tabbondexper}}
\begin{center}
    \begin{tabular}{ c c c c c c c}
    \hline
    Name & Bonded/DSP & Diameter & Thickness & Pattern Type & Pattern Depth & Notes \\ 
    - & - & mm & mm & - & nm & - \\ 
        \hline
    VG01 & Bonded  & 100  &$1$ $\times$ 2 & None  & 2/13/2013 &  \\
    VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  None  & 2/13/2013 &  \\
    VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & See Fig. \ref{figVG03pattern}\\    
    VG04 & Bonded & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15 & \\
    VG05 & DSP & 100 & 1&  - & - & -  \\
    VG06 & DSP & 100 & 3.5 &  - & - & - \\
    VG07 & - & - & - &  - & - & \\
    VG08 & - & - & - &  - & - & \\
    VG09 & Bonded & 75 & 1.5? & mesh C & 49 $\pm$6 & \\
    VG15 & Bonded & 100 & 1.5? &  mesh C, M, F  &14 $\pm$2 & \\
    VG17 & Bonded & 75 & 6.0? &  mesh C & 64 $\pm$6 nm & \\
    VS20 & Bonded & 100 & 1.5? &  mesh C, M, F & 95 $\pm$5 & \\
    \hline
    \end{tabular}
\end{center}
\end{table}

\section{Substrate preparation}
We cleaned the surfaces before bonding to minimize the interfacial particle density.  The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.  We measured the large scale surface flatness of the pucks.  The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.  We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.

The detailed bonding procedure will be chronicled in a future paper.  Briefly, we prepared the substrates with standard cleaning procedures of solvents in a megasonic.  We then applied MHz frequency oxygen plasma ashing.  We soaked the wafers in DI water then dried with an N$_2$ gun.  We pressed the wafers together from the center to the outside with normal human force.

@@

The air cavity is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.  Note that we imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.  The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.  We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.  We assume Si is nonabsorbing longward of $\lambda$ = 1150 nm.  We compute the reflectivity for an Si$-$air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.  The computation is simply Fresnel's law at normal incidence:

@@

The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.  We've defined the coefficient of Finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.  The coefficient of Finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.  In the next subsection we work out the wavelength dependent transmission for the air etalon

\subsection{Predicted transmission spectrum for bonded Si with finite interface gap}
\label{secTheory}
There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate.  Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.  The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.  We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.  Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.  The details are worked out in Appendix \ref{sec:Append-IMRTMM}.  We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with a gap thickness $d$.  The DSP Si puck with no air gap has a transmission equal to:

@@

which has an average value of about 53\%, as shown in the top curve of the Figure above.  Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.  If the Si wafer path length is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.  This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.
%We work out this behavior in the Appendix and show that in the limit $l>> \lambda^2 / \delta \lambda $ the spectral transmission tends to equation X.

@@

>>>>

While it is possible that proper precautions to eliminate all micron-sized particles from both Si surfaces prior to bonding, the power-law nature of the particle size distribution in clean-room environments \cite{doi:10.1080/02786828608959094} means that it is significantly more difficult to assure that no particles of the 10$-$60 nm size scale are present.  Small particles could be a concern if they act as "tent poles", that create gaps where the bonding is incomplete (Figure \ref{figParticle}).  Authors have referred to these large ($\sim$ 1 $\mu$m size) gaps as \emph{IR bubbles}, or simply bubbles, owing to their appearance in infrared (IR) transmission images as contiguous regions of a deficit of IR transmission.  They have also been referred to as voids, or interfacial voids, or gaps.  In this paper we will reserve the term ``IR bubbles'' for gaps that are large enough to be detected directly in conventional IR imaging.  The term ``gap'' will apply more broadly to any finite gap, which exceeds the scale over which interfacial molecular forces are acting, roughly say 2 nm.

@@

The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.  We classify the 3.3 mm thick substrates as ``pucks'' and the $\sim$1.0 mm thick substrates as ``wafers''.  The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.  There is also evidence that bond front propagation is slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.

@@

\begin{table}[h!]
\caption{UTexas Si bonding experiments \label{tabbondexper}}
\begin{center}
    \begin{tabular}{ c c c c c c c}
    \hline
    Name & Bonded/DSP & Diameter & Thickness & Pattern Type & Pattern Depth & Notes \\ 
    - & - & mm & mm & - & nm & - \\ 
        \hline
    VG01 & Bonded  & 100  &$1$ $\times$ 2 & None  & - &  \\
    VG02 & Bonded & 100 & $3.3$ $\times$ 2 &  None  & - &  \\
    VG03 & Bonded & 100 & $3.3$ $\times$ 2 &  Hole \&Petal & 4000$\pm$200 & See Fig. \ref{figVG03pattern}\\    
    VG04 & Bonded & 100 & $1$ $\times$ 2 &  Patchwork Islands & 15 & \\
    VG05 & DSP & 100 & 1&  - & - & Reference Sample  \\
    VG06 & DSP & 100 & 3.5 &  - & - & Reference Sample \\
    VG07 & - & - & - &  - & - & \\
    VG08 & - & - & - &  - & - & \\
    VG09 & Bonded & 75 & 1.5? & mesh C & 49 $\pm$6 & \\
    VG15 & Bonded & 100 & 1.5? &  mesh C, M, F  &14 $\pm$2 & \\
    VG17 & Bonded & 75 & 6.0? &  mesh C & 64 $\pm$6 nm & \\
    VS20 & Bonded & 100 & 1.5? &  mesh C, M, F & 95 $\pm$5 & \\
    \hline
    \end{tabular}
\end{center}
\end{table}

\section{Substrate preparation}
We cleaned the surfaces before bonding to minimize the interfacial particle density.  The surface roughness was typically about 2 nm, as measured with a Veeco NT9100 Optical Profiler.  We measured the large scale surface flatness of the pucks.  The Fisba2 interferometer has a 2 inch diameter beam at $\lambda=$632.8 nm.  We found a typical peak to valley surface flatness of 2 waves over the central 2 inch diameter.  We prepared the substrates with standard cleaning procedures of solvents in a megasonic.  We then applied MHz frequency oxygen plasma ashing.  We soaked the wafers in DI water then dried with an N$_2$ gun.  We pressed the wafers together from the center to the outside.

@@

The air cavity is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.  Note that we imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.  The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.  We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.  Si is negligibly absorbing longward of $\lambda$ = 1150 nm, which we verified with comparison of Si reference samples of different thicknesses.  We compute the reflectivity for an Si$-$air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.  In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ for clarity.  The computation is simply Fresnel's law at normal incidence:

@@

The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.  We've defined the coefficient of finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.  The coefficient of finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.  In the next subsection we work out the wavelength dependent transmission for the air etalon.

\subsection{Predicted transmission spectrum for bonded Si with finite interface gap}
\label{secTheory}
There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate at normal incidence.  Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.  The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.  We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.  Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.  The details are worked out in Appendix \ref{sec:Append-IMRTMM}.  We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with a gap thickness $d$.  The DSP Si puck with no air gap has a transmission equal to:

@@

which has an average value of about 53\%.  Note that this transmission is above a na\"ive value of $T_{DSP}=(1-R^2)$, which does not take into account multiple incoherent reflections, and is therefore an underestimate.  If the Si wafer path is less than the coherence length for the given spectral bandwidth, then the multiple reflections would interfere coherently, and the Si wafer would behave as a Fabry-P\`erot etalon.  This scenario is beyond the scope of this paper, since we are primarily interested in thick substrates with small gaps in between the interfaces.
%We work out this behavior in the Appendix and show that in the limit $l>> \lambda^2 / \delta \lambda $ the spectral transmission tends to equation X.

@@

\subsection{IR bubble detection and limitations}
